Do convolutional neural networks exhibit human-like attentional effects? Conflicting results from a visual search task

### 2019-05-27 docs/ccneuro2019/paper.tex
fix graphicspath and image names in paper.tex
add linebreaks in docs/ccneuro2019/paper.tex abstract
to be consistent with 90 char margin width
change paper title

\graphicspath{ {./images/} }

\title{What is attention? A study of neural networks performing a visual search task}

@@

\section{Abstract}
{
\bf
In machine learning and artificial intelligence, attention mechanisms are proposed to improve classification accuracy. In cognitive science, however, attention mechanisms are often defined in terms of experimental manipulations that impair performance. For example, psychologists have long used a now classic form of visual search task to determine what factors increase reaction times or decrease accuracy. Here we investigate 
the ability of neural networks to perform the classic visual search task, with the goal of reconciling what appear to be different definitions of 
attention in cognitive science and in artificial intelligence. We first show that convolutional neural networks (CNNs) trained for image classification 
can show human-like attentional effects, replicating and extending a previous study. Then we demonstrate that, by changing how we train the networks 
and the statistics of the training data, these effects can largely be eliminated, i.e., that CNNs are capable of performing the task without explicit 
attention mechanisms. We go on to show that manipulating the discriminability of the target and distractors in the task can produce attentional effects 
in spite of the changes we made to training. We discuss the implications of these findings for the study of attention in neuroscience and AI.
}
\begin{quote}
\small
\textbf{Keywords:} 
attention;visual search
\end{quote}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\columnwidth]{fig1}
\end{center}
\caption{A classic form of visual search task.} 
\label{sample-figure}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\columnwidth]{fig2}
\end{center}
\caption{CNNs trained for image classification show set size effects.} 
\label{sample-figure}
\end{figure}

>>>>

\graphicspath{ {../static/} }

\title{Do convolutional neural networks exhibit human-like attentional effects? 
Conflicting results from a visual search task}

@@

\section{Abstract}
{
\bf
In machine learning and artificial intelligence, attention mechanisms are proposed to 
improve classification accuracy. In cognitive science, however, attention mechanisms are 
often defined in terms of experimental manipulations that impair performance. For example,
psychologists have long used a now classic form of visual search task to determine what 
factors increase reaction times or decrease accuracy. Here we investigate the ability of 
neural networks to perform the classic visual search task, with the goal of reconciling 
what appear to be different definitions of attention in cognitive science and in 
artificial intelligence. We first show that convolutional neural networks (CNNs) trained 
for image classification can show human-like attentional effects, replicating and 
extending a previous study. Then we demonstrate that, by changing how we train the 
networks and the statistics of the training data, these effects can largely be 
eliminated, i.e., that CNNs are capable of performing the task without explicit attention 
mechanisms. We go on to show that manipulating the discriminability of the target and 
distractors in the task can produce attentional effects in spite of the changes we made to
training. We discuss the implications of these findings for the study of attention.
}
\begin{quote}
\small
\textbf{Keywords:} 
attention;visual search
\end{quote}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\columnwidth]{fig1/fig1.eps}
\end{center}
\caption{A classic form of visual search task.} 
\label{sample-figure}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\columnwidth]{fig2/fig2.png}
\end{center}
\caption{CNNs trained for image classification show set size effects.} 
\label{sample-figure}
\end{figure}
